Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-22 ICML2019/main.tex
working on the problem statement

\section{Preliminaries}
\label{sec_background}
\noindent \textbf{Bayesian inference.} 
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$ given $\dataobs$ is defined as:
\[
  \Pr(\theta | \dataobs) \equiv \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression $\Pr(\dataobs | \theta)$ denotes the
\emph{likelihood} of observing $\dataobs$ under a value of
$\theta$. Since we consider $\dataobs$ to be fixed, the likelihood is
a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$, that is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distribution of the prior. In this case we say that the prior
is conjugate to the likelihood function. Use of a conjugate prior
simplifies calculations and allows for inference to be performed in a
recursive fashion over the data.
% Then, we have:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
% \noindent \textbf{Beta-binomial System.}
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
We will consider the situation where the underlying data is drawn from a binomial distribution
($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
the parameter --informally called \emph{bias}-- of a Bernoulli
distributed random variable and $n$ is the number of draws. The
prior distribution over $\theta\in [0,1]$ is going to be a beta
distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
in section \ref{} when the private mechanisms will be described.
% \noindent \textbf{Dirichlet-multinomial Systems.}

The Beta-Binomial model can be immediately generalized to the Dirichlet-Multinomial model, with underlying data multinomially distributed.
The \emph{bias} is represented by the parameter $\vtheta$, the vector of parameters of a categorically distributed random variable. The prior distribution over $\vtheta\in [0,1]^{k}$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbb{R}^{+})^{k}$, with p.d.f:
\[
\Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
\Pr(\dataobs|\vtheta)\equiv\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where $\iverson{\cdot}$ represents the Iverson bracket notation.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots \Delta\alpha_k)$ the posterior distribution over $\vtheta$ turns out to be
\[
   \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)=\dirichlet{\valpha + \Delta \valpha} 
\]
where $+$ denotes the componentwise sum of vectors.

\noindent \textbf{Differential Privacy.}
\emph{Differential privacy} is a strong quantitative notion of statistical privacy proposed by \cite{dwork2006}.
Differential privacy of a mechanism implies that changing the input of a mechanism slightly only reflects
in a controlled and bounded way on the output distribution. 
Formally, the $\epsilon$ parameter
controls how much two outputs starting from \emph{close} inputs can differ. When two inputs are close we call them \emph{adjacent}.
By adjacency we usually mean a relation over input databases which includes all the pair of databases which differ by at most one row.
The formal definition of $\epsilon$-differential privacy follows.
\begin{definition}
\label{def_epsilon_dp}

@@

\noindent \textbf{Problem setting.}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in Section \ref{sec_background}.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta and Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the Beta distribution with 
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a private version of the Dirichlet distribution with parameter
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the values released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Section \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are the real and the perturbed pairs of parameters, respectively.
However, distance between
distributions are more naturally  measured by distribution
metrics such as $f$-divergences, e.g. Hellinger distance, etc., than the $\ell_1$ norm shown above.
{\color{blue} still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

@@

\section{Private Mechanisms}
\label{sec_mechs}
\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
We start off by presenting a mechanism based on the Laplace primitive.
The mechanism, denoted by $\lapmech$, takes in input the data $\dataobs$, a mechanism, $\bysinfer$, that is used to perform inference on some parameter,
a prior and a likelihood function, and the $\epsilon$ parameter measuring the privacy level.
Now, $\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$,
simply adds Laplace noise to the parameters of the posterior distributions. The mechanisms also performs a
post-processing phase by flooring and  truncating the noisy parameters so that they belong to the right interval.
In the case of the Beta-Binomial model the output produced is: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$
with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into
$[0,n]$. This mechanism is described by the pesudocode in Figure \ref{mech:lapmech}.\\
Since we are adding noise directly to the parameters, the sensitivity should be calibrated w.r.t $\ell_1$-norm. That is,
the sensitivity is defined as:
\[
   \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
\]
This value can be quickly seen as being upper bounded by  the dimensionality. (e.g. 2 for the Beta-Binomial model and  $k$ for
the Dirichlet-Multinomial model).
  \begin{algorithm}
  \caption{$\lapmech$ for the Beta-Binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad  $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad  $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad  $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ \\
  % \RETURN $a_{k+1}$.
  \noindent {\bf return} $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

@@

\subsection{A new family of mechanisms}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{??} where the scoring function is based on a metric over
probability distributoins.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta),\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS \equiv 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
%
  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $GS$ be the global sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

@@

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot \ux{r}}{2\cdot LS(\dataobs)})}
{\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)\equiv\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix.

>>>>

\section{Preliminaries}
\label{sec_background}
\noindent \textbf{Bayesian inference.} 
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$
given $\dataobs$ is defined by Bayes' Theorem as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression $\Pr(\dataobs | \theta)$ denotes the
\emph{likelihood} of observing $\dataobs$ under a value of
$\theta$. 
% Since we consider $\dataobs$ to be fixed, the likelihood is
% a function of $\theta$.
% For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
% Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
% the initial belief on $\theta$, that is, when no data has been
% observed. In practice though,
In parametric statistics, prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions as the prior. In this case we say that the prior
is conjugate to the likelihood function. Use of a conjugate prior
simplifies calculations and allows for inference to be performed in a
recursive fashion over the data. This approach is implemented in a
precise or approximate way in the compiler of several
probabilistic programming languages~\cite{}.

%  support bayesian inference
% by means of a language primitive, often called $\bysinfer$, which
% supports the (precise or approximate) parametric computation of the
% posterior of a given model. In the rest of the paper, we will use the following
% notation as a shorthand for the inference process:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
\noindent \textbf{Dirichlet-multinomial model.}
In this work we will focus on the Dirichlet-Multinomial model for
categorical data and its restriction to binary data: the Beta-Binomial
model. 
That is, we will consider the situation where 
the underlying categorical data is drawn from a multinomial
distribution with parameter the vector $\vtheta\in [0,1]^{k}$. The prior distribution over $\vtheta$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for,
$\valpha\in(\mathbb{R}^{+})^{k}$ and $k\in\mathbb{N}$, with p.d.f:
\[
\Pr(\vtheta)=\frac{1}{\mbetaf(\valpha)}\cdot\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ is a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $|\datauni|=k$.
The likelihood function is
$
\Pr(\dataobs|\vtheta)=\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
$
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\mathbf{1}_{\iverson{x_j=a_i}}$.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots
\Delta\alpha_k)$ the posterior distribution over $\vtheta$ is
$\dirichlet{\valpha + \Delta \valpha}$, where $+$ denotes the componentwise sum of vectors.
In the following we will use the notation $\bysdir(n,\vtheta, \valpha,
\dataobs)$ to denote the posterior of this process.
%i.e., $\bysdir(n,\vtheta, \valpha, \dataobs)=\dirichlet{\valpha + \Delta \valpha}$.
%
% ($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
% the parameter --informally called \emph{bias}-- of a Bernoulli
% distributed random variable and $n$ is the number of draws. The
% prior distribution over $\theta\in [0,1]$ is going to be a beta
% distribution, $\betad(\alpha, \beta)$, with parameters
% $\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
% \[
%   \Pr(\theta)= \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
% \]
% where $\betaf(\cdot,\cdot)$ is the beta function.
% The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
% \[
%   \Pr(\dataobs | \theta)= \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
% \]
% where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
% The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
% From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
% posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
% in section \ref{} when the private mechanisms will be described.

% We will consider the situation where the underlying data is drawn from a binomial distribution
% ($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
% the parameter --informally called \emph{bias}-- of a Bernoulli
% distributed random variable and $n$ is the number of draws. The
% prior distribution over $\theta\in [0,1]$ is going to be a beta
% distribution, $\betad(\alpha, \beta)$, with parameters
% $\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
% \[
%   \Pr(\theta)= \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
% \]
% where $\betaf(\cdot,\cdot)$ is the beta function.
% The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
% \[
%   \Pr(\dataobs | \theta)= \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
% \]
% where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
% The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
% From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
% posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
% in section \ref{} when the private mechanisms will be described.
% \noindent \textbf{Dirichlet-multinomial Systems.}
%
The Beta-Binomial model can be obtained from the Dirichlet-Multinomial
model by fixing $k=2$.

\noindent \textbf{Differential Privacy.}
\emph{Differential privacy}~\cite{dwork2006} is a quantitative notion
of data privacy guaranteeing  a bound on the that changing the input of a mechanism slightly only reflects
in a controlled and bounded way on the output distribution. 
Formally, the $\epsilon$ parameter
controls how much two outputs starting from \emph{close} inputs can differ. When two inputs are close we call them \emph{adjacent}.
By adjacency we usually mean a relation over input databases which includes all the pair of databases which differ by at most one row.
The formal definition of $\epsilon$-differential privacy follows.
\begin{definition}
\label{def_epsilon_dp}

@@

\section{Problem Statement}
\label{subsec_problem}
We are interested in exploring the design space of mechanisms for privately releasing the posterior
distribution for the Dirichlet-Multinomial model as described in the
previous section.
It is worth noticing that the posterior distribution of this model is fully characterized
by its parameters. That is, we are interested in releasing a private version of
the Dirichlet distribution with parameter
$\valpha'=(\valpha + \Delta \valpha)$, or in the case of the
Beta-Binomial model we are interested in releasing a private version of the Beta distribution with 
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n -
\Delta \alpha)$.
\citet{zhang2016differential} and \citet{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the values released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Section \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are the real and the perturbed pairs of parameters, respectively.
However, distance between
distributions are more naturally  measured by distribution
metrics such as $f$-divergences, e.g. Hellinger distance, etc., than the $\ell_1$ norm shown above.
{\color{blue} still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

@@

\section{Private Mechanisms}
\label{sec_mechs}
\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
We start off by presenting a mechanism based on the Laplace primitive.
The mechanism, denoted by $\lapmech$, takes in input the data $\dataobs$, a mechanism, $\bysinfer$, that is used to perform inference on some parameter,
a prior and a likelihood function, and the $\epsilon$ parameter measuring the privacy level.
Now, $\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$,
simply adds Laplace noise to the parameters of the posterior distributions. The mechanisms also performs a
post-processing phase by flooring and  truncating the noisy parameters so that they belong to the right interval.
In the case of the Beta-Binomial model the output produced is: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$
with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into
$[0,n]$. This mechanism is described by the pesudocode in Figure \ref{mech:lapmech}.\\
Since we are adding noise directly to the parameters, the sensitivity should be calibrated w.r.t $\ell_1$-norm. That is,
the sensitivity is defined as:
\[
   \Delta \bysinfer =\max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
\]
This value can be quickly seen as being upper bounded by  the dimensionality. (e.g. 2 for the Beta-Binomial model and  $k$ for
the Dirichlet-Multinomial model).
  \begin{algorithm}
  \caption{$\lapmech$ for the Beta-Binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad  $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad  $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad  $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ \\
  % \RETURN $a_{k+1}$.
  \noindent {\bf return} $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

@@

\subsection{A new family of mechanisms}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{??} where the scoring function is based on a metric over
probability distributoins.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset=\{\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta),\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS = 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
%
  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $GS$ be the global sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

@@

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot \ux{r}}{2\cdot LS(\dataobs)})}
{\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)=\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix.
